```{r knitr-setup, warning=FALSE, message=FALSE, error=FALSE, echo=FALSE, results='hide'}

library(knitr)
opts_chunk$set(warning = FALSE,
               message = FALSE,
               error = FALSE,
               cache=TRUE,
               results = "hide",
               echo=opts_knit$get("rmarkdown.pandoc.to") != 'latex')

options(digits=2)

## Produce markdown-formatted figures so that pandoc knows what to do with
## the captions. requires pandoc-fignos to parse the IDs. refer to figures
## in text with {@fig:label} or just @fig:label
## 
## (see https://github.com/tomduck/pandoc-fignos)
knit_hooks$set(plot = function(x, options) {
  paste0('![', options$fig.cap, ']',
         '(', opts_knit$get('base.url'), paste(x, collapse='.'), ')',
         '{#fig:', options$label, '}')
})
## Produce markdown-formatted table captions with anchors for cross-refs.
## Requires pandoc-tablenos to parse the IDs. Refer to tables
## in text with {@tbl:label} or @tbl:label.
## Based partly on http://stackoverflow.com/a/18672268
##
## (see https://github.com/tomduck/pandoc-tablenos)
knit_hooks$set(tbl.cap = function(before, options, envir) {
  if(!before){
    paste0('\n\nTable: ', options$tbl.cap,
           ' {#tbl:', options$label, '}', sep = '')
  }
})

```

```{r preamble, cache=FALSE}

library(tidyverse)
library(dplyr)
library(glue)
library(magrittr)
library(beliefupdatr)
library(supunsup)

library(rstan)
library(brms)
library(tidybayes)

library(cowplot)
theme_set(cowplot::theme_cowplot() +
            theme(plot.title=element_text(hjust=0)))
library(latex2exp)

## devtools::install_github('kleinschmidt/daver')
library(daver)
## devtools::install_github('kleinschmidt/phonetic-sup-unsup')
library(supunsup)
## devtools::install_github('kleinschmidt/beliefupdatr')
library(beliefupdatr)

```


# Experiment 1

The Bayesian logistic regression model for Experiment 1 was fit using `brms`
[@Burkner2017], based on the formula:

```
respP ~ 1 + bvotCond * vot_s * trial_s + (1 + vot_s | subject)
```

`respP` is 0 for a click on the /b/-initial response picture and 1 for the
/p/-initial, `bvotCond` is the experimental condition (exposure distributions),
dummy coded as 0 or 1, and `vot_s` and `trial_s` are the stimulus VOT and the
trial number, each standardized to have mean 0 and variance 1.

This formula generates fixed effects for all main effects of condition, VOT, and
trial, all two-way interactions, and the three-way interaction.

```{r expt-1-model-table, results="raw"}

b_logit_exp1 <- readRDS("models/brm_logistic_exp1.rds")
summary(b_logit_exp1)

```

```{r expt-1-model-plot, fig.width=8.9, fig.height=6.4, fig.cap="Posterior distributions for fixed-effects coefficients in Experiment 1.  Red indicates coefficients that are reliably less than zero, and blue indicates reliably greater than zero.}

b_logit_exp1 %>%
  gather_draws(`b_.*`, regex=TRUE) %>%
  ungroup() %>%
  mutate(.variable = factor(str_replace(.variable, "b_", ""),
                            levels = rownames(fixef(b_logit_exp1)))) %>%
  group_by(.variable) %>%
  mutate(p_gt_zero = mean(.value > 0)) %>%
  ggplot(aes(x=.value, y=.variable, fill=p_gt_zero)) +
  geom_halfeyeh(relative_scale = 2) +
  scale_fill_gradient2("p(β > 0)", midpoint=0.5, mid=gray(0.9)) +
  geom_vline(xintercept = 0, linetype=3)

```
